Symmetric Circuits for Rank Gates

### 2018-01-24 sections/translating-formulas-to-cicuits.tex
24th January 2018

\documentclass[../paper.tex]{subfiles}
\begin{document}

\subsection{Algorithms for Circuits}
In order to translate between $P$-uniform families of circuits and formulas of
$\FPR$ we will require that certain important circuit transformations can be
completed in polynomial time. In this subsection we define a number of
operations that perform commonly invoked operations on circuits, and which can
easily be shown to be implementable as algorithms that run in polynomial time.
In each case we assume a reasonable encoding of the circuit.

We first define the operation $\op{merge}$ that takes in a circuit and a set of
syntactically equivalent gates and produces a new circuit by combining those
gates into a single gate.

\begin{definition}
  The operation $\op{merge}(C, H)$ takes as input a circuit $C = \langle G,
  \Omega, \Sigma , \Lambda, L \rangle$ and a set of syntactically equivalent
  gates $H := \{h_1, \ldots , h_p\}$ and outputs the circuit $C' = \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed by an algorithm running in
time polynomial in the size of the input circuit. We note that the input and
output circuits compute the same function. Moreover, if the input circuit is
symmetric (resp. has unique labels, is transparent) then the output circuit will
also be symmetric (resp. have unique labels, is transparent).

The the operation $\op{merge-all}$ simply applies $\op{merge}$ to the syntactic
equivalence classes of some input set of gates.

\begin{definition}
  The operation $\op{merge-all}(C, H)$ takes as input an ordered circuit $C$ and
  a set of gates $H$ and outputs another circuit $C'$. Let $H_1, \ldots, H_p$ be
  the syntactic equivalence classes of $H$ linearly ordered using the order on
  the circuit but such that for all $i, j \in [p]$ if $i \leq j$ then the height
  of the gates in $H_i$ is less than or equal to the height of the gates in
  $H_j$. Let $C'$ be the result of successively executing $\op{merge}$ on each
  equivalence class.
\end{definition}

For an input tuple $(C, H)$ if $H$ is a set of gates with unique labels it
follows from Lemma \cite{} that the syntactic equivalence relation on $H$ can be
computed in time polynomial in the size of $C$. Thus, if the set $H$ has unique
labels it can be shown that the operation $\op{merge-all}$ can be computed in
time polynomial in the size of the input circuit. As in the case of
$\op{merge}$, the operation $\op{merge-all}$ preserves symmetry, unique labels
and the ensures that the input and output circuits compute the same function. If
$C$ is an ordered circuit and $H$ is a set of gates with unique labels, then if
$g$ is a gate in both $C$ and $C' := \op{merge-all}(C, H)$, then if all of the
input gates of $g$ in $C$ are in $H$ then no two distinct input gates of $g$ in
$C'$ are syntactically equivalent (but $g$ does not have unique labels in $C'$
as $L'(g)$ may not be injective).

@@

The operation $\op{make-injective}$ takes in a circuit and a symmetric gate and
outputs an equivalent circuit but such that the selected gate now has injective
labels.

\begin{definition}
  The operation $\op{make-injective}(C, h)$ takes as input an ordered
  $(\mathbb{B}, \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L
  \rangle$ and a gate $h \in G$ and outputs the circuit $C' := \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$. Let $c = \max \{\vert
  L(g)^{-1}(h)\vert : g \in W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c >
  1$ then $C'$ is defined such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemize}
\end{definition}

@@

The operation $\op{make-injective-all}$ takes in a circuit and outputs an
equivalent circuit such that every symmetric-gate in that circuit has injective
labels.

\begin{definition}
  The operation $\op{make-injective-all}(C)$ takes as input an ordered
  $(\mathbb{B}, \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L
  \rangle$ and outputs the circuit $C' := \langle G', \Omega', \Sigma',
  \Lambda', L' \rangle$ defined as follows. Let $F(C*) = \op{make-injective}(C*,
  h)$, where $h$ is the first gate in $C*$ such that there exists a gate $g$ in
  $C*$ where $\vert L*(g)^{-1}(h)\vert > 1$. Let $(C^0, C^1, \ldots)$ be a
  sequence of circuits where $C^0 := C$ and for $i > 1$, $C^i = F(C^{i-1})$. It
  can be shown that this sequence converges after at most $\vert C \vert$ many
  steps. Let $C'$ be this limit.
\end{definition}

Since $\op{make-injective}$ can be computed in polynomial time, and the number
of iterations required in order to reach is a fixed-point in the definition of
$\op{make-injective-all}$ is bounded by the size of the input circuit, it
follows that $\op{make-injective-all}$ can be computed in time polynomial in the
size of the input circuit. Moreover, $\op{make-injective-all}$ preserves
symmetry, rigidity and unique labels. Moreover, the input and output circuits
compute the same function.

@@

The operation $\op{insert}$ takes as input two circuits and outputs a circuit
formed by inserting the gates of the first circuit into the second circuit.

\begin{definition}
  The operation $\op{insert}(C^1, C^2)$ takes as input a $(\mathbb{B}_1,
  \tau_1)$-circuit $C^1 := \langle G_1, \Omega_1, \Sigma_1 , \Lambda_1, L_1
  \rangle$ and a $(\mathbb{B}_2, \tau_2)$-circuit $C^2 := \langle G_2, \Omega_2,
  \Sigma_2 , \Lambda_2, L_2 \rangle$ and outputs a $(\mathbb{B}_1 \cup
  \mathbb{B}_2, \tau_1 \uplus \tau_2)$-circuit $C = \langle G, \Omega, \Sigma ,
  \Lambda, L \rangle$ where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G = G_1 \uplus G_2$,
  \item $\Omega = \Omega_2$,
  \item $\Sigma = \Sigma_1 \uplus \Sigma_2$,
  \item $\Lambda = \Lambda_1 \uplus \Lambda_2$, and
  \item $L = L_1 \uplus L_2$.
  \end{itemize}
\end{definition}

It can be shown that $\op{insert}$ can be be computed in polynomial time. If the
input circuits are symmetric then the output circuit will be symmetric.

The operation $\op{replace}$ removes a gate from a circuit and replaces it with
another gate.

\begin{definition}
  The operation $\op{replace}(C, g_1, g_2)$ takes as input a circuit $C =
  \langle G, \Omega, \Sigma, \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$
  in $C$ and outputs a circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$ such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and if then $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that the operation $\op{replace}$ can be computed in time
polynomial in the size of the input circuit.

The operation $\op{tag}$ takes in a circuit, a gate in that circuit, a relation
symbol and a tuple and outputs a circuit computing the same function but with an
additional gadget, constructed using the gates labelled by the given relation
symbol and tuple, marking the designated gate. The idea is that marking a gate
can be used as a step to ensure unique labels in a circuit.

\begin{definition}
  The operation $\op{tag}(C, R, g, \vec{a})$ takes in a $(\mathbb{B},
  \tau)$-circuit $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a
  relation symbol $R \in \tau$, a gate $g \in G$ and a tuple $\vec{a} \in [n]^k$
  for some $k$ and outputs the circuit $C_n' = \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ defined by the following algorithm. Let $G' = G \uplus
  \{\land_g, \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and
  $\Lambda' = \Lambda$. Define $L'$ and $\Sigma'$ as follows. For each $g \in G$
  $\Sigma'(g) = \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots , \land_k
  \}$, $\Sigma(g)$ is an $\AND$-gate that takes in two inputs. $\Sigma(\land_g)$
  is an $\AND$-gate that takes in a single input. Set $g$ to be the lone input
  to $\land_g$. Set the two inputs to $\land_0$ to be $\land_g$ and $\land_1$.
  For each $i \in [k-1]$ set the inputs of $\land_i$ to be $\land_{i+1}$ and the
  gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and $\Lambda^{-1}(g_i) =
  \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple $(\vec{a}(i), \ldots,
  \vec{a}(i))$. Set the inputs of $\land_k$ to be the constant gate $1$ and the
  gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) = \vec{a}_k$.

  For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$, if
  $L(h)(\vec{b}) = g$ then let $L'(h)(\vec{b}) = \land_0$, otherwise let
  $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

@@

It can be shown that $\op{concatenate}$ can be implemented so as to run in time
polynomial in the size of the input. The input and output circuits compute the
same function. Moreover, if the input circuit had unique labels then the output
circuit will have unique labels. In the event that the input relation symbol is
obvious from the context we omit it when invoking this algorithm.

The operation $\op{concatenate}$ takes in a circuit and a set of appropriate
circuits and forms a new circuit by replacing the relational gates of the
designated circuit with the output gates of the circuits in the given set of
circuits.

\begin{definition}
  The operation $\op{concatenate}(C, \mathcal{C}, \tau_R, R)$, where $C :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$ is a $(\mathbb{B},
  \tau)$-circuit, $\tau_R := (R_1, \ldots, R_t)$ is a sequence of relation
  symbols such that $\tau_R \subseteq \tau$, $\mathcal{C} := ( C^1, \ldots,
  C^t)$ is a finite sequence of circuits, where for each $i \in [t]$, $C^i =
  \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a $(\mathbb{B}_i,
  \tau_i)$-circuit such that $\tau_i \not\subset \tau_R$ and $C^i$ computes a
  query of arity $\arty(R_i)$, and $R$ is a relation symbol in $(\tau \setminus
  \tau_R) \cup (\bigcup_{1 \leq i \leq t},\tau_i))$. The operation outputs a
  circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined from
  the input as follows. Note that all circuits take in structures of size $n$.

  Let ${C^0}' := \langle G^0, \Omega^0, \Sigma^0, \Lambda^0, L^0 \rangle := C$.
  For each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus \tau_R$.
  For each $i \in [t]$ construct the circuit ${C^{i}}' = \langle G_i',
  \Omega_i', \Sigma_i', \Lambda_i', L_i' \rangle $ by $\op{inserting}$ $C^i$
  into ${C^{i-1}}'$, and call the resultant circuit ${D^{i-1}}'$. For each
  relational gate $g$ in $D^{i-1}$ labelled by the symbol $R_i$, $\op{replace}$
  $g$ with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ and $\op{tag}$
  the gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
  $\Lambda_{i-1}'(g)$ and the relation symbol $R$. Output the circuit $C' :=
  {C^t}'$.
\end{definition}

@@

It can be shown that $\op(concatenate)$ can be computed in time polynomial in
the sum of the sizes of the input circuits. Moreover, if the input circuits are
all symmetric then the output circuit will be symmetric. Furthermore, if the
input sequences of circuits and relations have just a single element, then if
the input circuits have unique labels then the output circuit will have unique
labels.

\subsection{Transparency and Unique Extensions}
\label{sec:transparency-and-unique-extensions}
% In contrast with the work of Anderson and Dawar~\cite{AndersonD17}, the
% circuit families we use to establish the main result of this paper are
% additionally required to be transparent. In this subsection we briefly discuss
% this condition, and present results suggesting its necessity.

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}
In this section we first introduce a number of useful polynomial-time computable
functions which define transformations on circuits. In the second subsection we
use these these functions to define a transformation of $\FPR$-formulas into
equivalent families of transparent rank-circuits.

\subsection{Algorithms for Circuits}
In order to simplify the presentation of the translation between $P$-uniform
families of circuits and formulas of $\FPR$, as well as the proofs of other
results (see Lemma~\ref{lem:transparent-unique}), we define a number of
often-used polynomial-time computable circuit transformations.

We first define the function $\op{merge}$, which maps a circuit and a set of
syntactically-equivalent gates to a circuit formed by combining those gates into
a single gate.

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed in time polynomial in the size
of the input circuit. We note that the input and output circuits compute the
same function. Moreover, if the input circuit is symmetric (resp. has unique
labels, is transparent) then the output circuit will also be symmetric (resp.
have unique labels, is transparent).

We now define the function $\op{merge-all}$ which maps a circuit and a set of
gates to a new circuit formed by applying $\op{merge}$ to the
syntactic-equivalence classes of the set of gates.

\begin{definition}
  \label{def:merge-all}
  The function $\op{merge-all}$ maps a circuit $C$ and a set of gates $H$ to a
  circuit $\op{merge-all}(C, H) := C'$, where $C'$ is defined as follows. Let
  $H_1, \ldots, H_p$ be the syntactic equivalence classes of $H$, and let $C'$
  be defined from $C$ by successively applying $\op{merge}$ to each of these
  equivalence classes.
\end{definition}

We can compute $\op{merge-all}$ using the following algorithm. Let $(C, H)$ be a
tuple in the domain of $\op{merge-all}$. We assume a reasonable encoding of the
circuit, and note that any such encoding induces a linear order on the
structure. Let $H_1, \ldots, H_p$ be the syntactic equivalence classes of $H$
linearly ordered such that for all $i, j \in [p]$ if $i \leq j$ the height of
the gates in $H_i$ is less than or equal to the height of the gates in $H_j$.
Let $C'$ be the result of successively executing the algorithm for $\op{merge}$
on each such equivalence class in order.

It is worth noting that for an input tuple $(C, H)$, if follows from
Lemma~\cite{lem:unique-labels-syntactic-equiv} that if $H$ has unique labels
then the syntactic equivalence relation on $H$ can be computed in time
polynomial in the size of $C$. Thus, if the set $H$ has unique labels, it can be
shown that the operation $\op{merge-all}$ can be computed in time polynomial in
the size of the input circuit.

As in the case of $\op{merge}$, the function $\op{merge-all}$ preserves
symmetry, unique labels and the ensures that the input and output circuits
compute the same function. Let be a circuit $C$ a circuit and $H$ a set of gates
in $C$ with unique labels. Then if $g$ is a gate in both $C$ and $C' :=
\op{merge-all}(C, H)$, and all of the children of $g$ in $C$ are in $H$, then no
two distinct children of $g$ in $C'$ are syntactically-equivalent. We note we
have no guarantee that $g$ will have unique labels in $C'$ as $L'(g)$ may not be
injective.

@@

The now define the function $\op{make-injective}$ that maps a circuit and a
symmetric gate to an equivalent circuit containing all the same gates as the
input circuit but such that the specified gate now has injective labels.

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemize}
\end{definition}

@@

We now define the function $\op{make-injective-all}$ that maps a circuit to an
equivalent circuit such that every symmetric gate now has injective labels.

\begin{definition}
  The function $\op{make-injective-all}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ to a circuit $\op{make-injective}(C) := C' :=
  \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined as follows.

  Let $F$ be a function that maps an ordered circuit $C*$ to the circuit
  $\op{make-injective}(C*, h)$, where $h$ is the topologically first gate such
  that there exists a gate $g$ in $C*$ where $h$ is a child of $g$ and $\vert
  L*(g)^{-1}(h) \vert > 1$. Let $(C^0, C^1, \ldots)$ be a sequence of circuits
  where $C^0 := C$ and for $i > 1$, $C^i := F(C^{i-1})$. It can be shown that
  this sequence converges after at most $\vert C \vert$ many steps. Let $C'$ be
  the limit of this sequence.
\end{definition}

Since $\op{make-injective}$ can be computed in polynomial time, and the number
of iterations required to converge to the limit in the above definition is
bounded by the size of the circuit, it can be shown that
$\op{make-injective-all}$ can be computed in time polynomial in the size of the
input circuit. Furthermore, we note that $\op{make-injective-all}$ preserves
symmetry, rigidity and unique labels. Moreover, the input and output circuits
compute the same function.

@@

We now define the function $\op{insert}$ that maps two circuits to a circuit
formed by inserting the gates of the first circuit into the second circuit.

\begin{definition}
  The function $\op{insert}$ maps a $(\mathbb{B}_1, \tau_1)$-circuit $C^1 :=
  \langle G_1, \Omega_1, \Sigma_1 , \Lambda_1, L_1 \rangle$ and a
  $(\mathbb{B}_2, \tau_2)$-circuit $C^2 := \langle G_2, \Omega_2, \Sigma_2 ,
  \Lambda_2, L_2 \rangle$ to a $(\mathbb{B}_1 \cup \mathbb{B}_2, \tau_1 \uplus
  \tau_2)$-circuit $\op{insert}(C^1, C^2) := C := \langle G, \Omega, \Sigma ,
  \Lambda, L \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G = G_1 \uplus G_2$,
  \item $\Omega = \Omega_2$,
  \item $\Sigma = \Sigma_1 \uplus \Sigma_2$,
  \item $\Lambda = \Lambda_1 \uplus \Lambda_2$, and
  \item $L = L_1 \uplus L_2$.
  \end{itemize}
\end{definition}

It can be shown that $\op{insert}$ can be be computed in time polynomial in the
size of the input circuits. Moreover, if the input circuits are symmetric then
the output circuit will be symmetric.

We now define the function $\op{replace}$ that maps a circuit $C$ and a pair of
gates $(g_1, g_2)$ in $C$ to a circuit $C'$ formed by replacing $g_1$ with $g_2$
in $C$.

\begin{definition}
  The function $\op{replace}$ maps a circuit $C = \langle G, \Omega, \Sigma,
  \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$ in $C$ to a circuit
  $\op{replace}(C, g_1, g_2) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and otherwise $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that the function $\op{replace}$ can be computed in time
polynomial in the size of the input circuit.

We now define the function $\op{tag}$ that maps a circuit, a relation symbol, a
gate in that circuit, and a tuple to a circuit computing the same function as
the input circuit but with an additional gadget, constructed using the gates
labelled by the given relation symbol and tuple, marking the designated gate.
The intuition is that this marking helps to distinguish a gate, and is used as a
step to ensure that a circuit has unique labels.

\begin{definition}
  The function $\op{tag}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle G,
  \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a gate $g \in G$, and a tuple $\vec{a} \in [n]^k$ for some $k$ to a
  circuit $\op{tag}(C, R, g, \vec{a}) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$, such that $G' = G \uplus \{\land_g,
  \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda' =
  \Lambda$. For each $g \in G$ $\Sigma'(g) = \Sigma(g)$. For each $g \in
  \{\land_0, \land_1, \ldots , \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that
  takes in two inputs. $\Sigma(\land_g)$ is an $\AND$-gate that takes in a
  single input, $g$ is the lone input to $\land_g$, and $\land_0$ has as input
  $\land_g$ and $\land_1$. For each $i \in [k-1]$ the inputs of $\land_i$ are
  $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$ and
  $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
  $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs of $\land_k$ are the constant
  gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and $\Lambda^{-1}(g_k) =
  \vec{a}_k$. For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$,
  if $L(h)(\vec{b}) = g$ then $L'(h)(\vec{b}) = \land_0$, otherwise
  $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

@@

It can be shown that $\op{tag}$ can be computed in time polynomial in the size
of its input. The input and output circuits compute the same function. Moreover,
if the input circuit has unique labels then the output circuit will have unique
labels. If the input relation symbol is obvious from the context we omit it when
invoking $\op{tag}$.

\begin{definition}
  The function $\op{tag-all}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle
  G, \Omega, \Sigma, \Lambda, L \rangle$ of order $n$, a relation symbol $R \in
  \tau$, a set of gates $H \subseteq G$, and $T := (\vec{a}_h)_{h \in H}
  \subseteq [n]^{k}$ a sequence of tuples indexed by $H$ for some fixed $k$, to
  a circuit $\op{tag-all}(C, R, H, T) := C' := \langle G', \Omega', \Sigma',
  \Lambda', L'\rangle$ of order $n$ defined as follows.

  Suppose there is a linear order on $H$, and write $H = \{h_1, \ldots, h_t\}$.
  Let $C_0 := C$ and for all $i \in [t]$ let $C_i := \op{tag}(C_{i-1}, R, h_i,
  T(h_i))$. Then $C' = C_{t}$.

  % such that $G' = G \uplus \{\land_g, \land_{g'}, \land_1, \ldots ,
  % \land_k\}$,
  % $\Omega' = \Omega$, and $\Lambda' = \Lambda$. For each $g \in G$ $\Sigma'(g)
  % =
  % \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots , \land_k \}$,
  % $\Sigma(g)$ is an $\AND$-gate that takes in two inputs. $\Sigma(\land_g)$ is
  % an $\AND$-gate that takes in a single input, $g$ is the lone input to
  % $\land_g$, and $\land_0$ has as input $\land_g$ and $\land_1$. For each $i
  % \in
  % [k-1]$ the inputs of $\land_i$ are $\land_{i+1}$ and the gate $g_i \in G'$
  % such that $\Sigma(g_i) = R$ and $\Lambda^{-1}(g_i) = \vec{a}_i$, where
  % $\vec{a}_i$ is the $k$-tuple $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs
  % of
  % $\land_k$ are the constant gate $1$ and the gate $g_k$ such that $\Sigma(g)
  % =
  % R$ and $\Lambda^{-1}(g_k) = \vec{a}_k$. For each gate $h \in G \setminus
  % \{g\}$ and $\vec{b} \in \ind(h)$, if $L(h)(\vec{b}) = g$ then
  % $L'(h)(\vec{b})
  % = \land_0$, otherwise $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

It is easy to see that, since $\op{tag}$ can be computed in time polynomial in
the combined size of its inputs, $\op{tag-all}$ can also be computed in time
polynomial in the combined size of its inputs.

We now define the function $\op{compose}$ that maps a circuit and a sequence of
appropriate circuits and to a circuit formed by replacing the input gates of the
designated circuit with the output gates of the circuits in the sequence of
circuits.

\begin{definition}
  The function $\op{compose}$ maps a $(\mathbb{B}, \tau)$-circuit $C := \langle
  G, \Omega, \Sigma, \Lambda, L \rangle$, a sequence of relation sybmols $\tau_R
  := (R_1, \ldots, R_t)$ such that $\tau_R \subseteq \tau$, a sequence of
  circuits $\mathcal{C} := ( C^1, \ldots, C^t)$, where for each $i \in [t]$,
  $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
  $(\mathbb{B}_i, \tau_i)$-circuit, and $\tau_i \not\subset \tau_R$ and $C^i$
  computes a query of arity $\arty(R_i)$ to a circuit $\op{compose}(C,
  \mathcal{C}, \tau_R) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$. Note that all circuits take in structures of size $n$. The circuit
  $C'$ is defined as follows.

  Let ${C^0}' := \langle G_0', \Omega_0', \Sigma_0', \Lambda_0', L_0' \rangle :=
  C$, and let $i \in [t]$. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$ and
  $G^{i-1}_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by
  the relational symbol in $R_i$. Let ${C^{i-1}}' := \langle G_i', \Omega_i',
  \Sigma_i', \Lambda_i', L_i' \rangle$ be the circuit formed from $D^{i-1}$ by
  replacing each gate in $g \in G^{i-1}_{R_i}$ with the gate
  $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
  $op{replace}$ function). Let $C' = {C^{t}}'$.
\end{definition}

% \begin{definition}
%   The function $\op{concatenate}$ maps a $(\mathbb{B}, \tau)$-circuit $C :=
%   \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a sequence of relation
%   sybmols $\tau_R := (R_1, \ldots, R_t)$ such that $\tau_R \subseteq \tau$, a
%   sequence of circuits $\mathcal{C} := ( C^1, \ldots, C^t)$, where for each $i
%   \in [t]$, $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is
%   a $(\mathbb{B}_i, \tau_i)$-circuit, $\tau_i \not\subset \tau_R$ and $C^i$
%   computes a query of arity $\arty(R_i)$, and $R$ is a relation symbol in
%   $(\tau \setminus \tau_R) \cup (\bigcup_{1 \leq i \leq t},\tau_i))$ to a
%   circuit $\op{concatenate}(C, \mathcal{C}, \tau_R, R) := C' := \langle G',
%   \Omega', \Sigma', \Lambda', L' \rangle$. Note that all circuits take in
%   structures of size $n$. The circuit $C'$ is defined as follows.

%   Let ${C^0}' := \langle G^0, \Omega^0, \Sigma^0, \Lambda^0, L^0 \rangle :=
%   C$. For each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus
%   \tau_R$. For each $i \in [t]$ let ${C^{i}}' := \langle G_i', \Omega_i',
%   \Sigma_i', \Lambda_i', L_i' \rangle := \op{insert} (C^i, {C^{i-1}}')$ be
%   defined as follows. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$. Let
%   $G_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by the
%   relational symbol in $R_i$. Let ${C^{i}}'$ be the circuit formed by
%   replacing each gate in $g \in G_{R_i}$ with the gate
%   $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
%   $op{replace}$ function) and using the $\op{tag}$ function on each gate
%   $\Omega_i (\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
%   $\Lambda_{i-1}'(g)$ and the relation symbol $V_i$. We then let $C' :=
%   {C^{t}}'$.

%   %   For each relational gate $g$ in $D^{i-1}$ labelled by the symbol
%   %   $R_i$,
%   %   $\op{replace}$ $g$ with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in
%   %   $D^{i-1}$
%   %   and $\op{tag}$ the gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$
%   %   using
%   %   the
%   %   tuple $\Lambda_{i-1}'(g)$ and the relation symbol $R$. Output the
%   %   circuit
%   %   $C'
%   %   := {C^t}'$.
% \end{definition}
% GW: do this tomorrow.

@@

It can be shown that $\op{compose}$ can be computed in time polynomial in the
sum of the sizes of the input circuits. Moreover, if the input circuits are all
symmetric then the output circuit will be symmetric. Also, for an input tuple
$(C, \mathcal{C}, \tau_R)$, we have that if $C$ has symmetric gates and
transparent, and all the circuits in $\mathcal{C}$ are transparent, then $C' :=
\op{compose}(C, \mathcal{C}, \tau_R)$ is transparent. If $C$ does not have
symmetric gates then, as the output gates of the circuits in $\mathcal{C}$ are
no longer labelled as output gates in $C'$, it may be that some of these gates
are syntactically-equivalent in $C'$, resulting in $C'$ not being transparent.

In certain contexts we are interested composing gates such that if the input circuits are
transparent (resp. have unique labels) then the output circuits will be
transparent (resp. have unique labels). In this case, while composing the
circuits we could apply the $\op{tag}$ function to the output gates of the
circuits in the input sequence. We now define the function
$\op{compose-and-tag}$ that implements this mapping.
